---
title: "PS1_Markdown_Trevisani"
author: 
- José Maria de Mattamouros Resende Fonseca de Oliveira (22-602-783)
- Linda Fiorina Odermatt (17-946-310)
- Elena Trevisani (22-620-603) 
- Vasily Zhuravlev (18-502-401) 
date: "2022-11-28"
output:
  pdf_document: default
  html_document: default
---




## Exercise 1

```{r, results=FALSE}
load("/Users/nenetrevisani/Desktop/Data Analysis 1/DA1/insurance-all.Rdata")
```
```{r}
nrow(data)
ncol(data)
```
observations = 1204;
colums of data= 7. 

As we mostly collect data to examine the dependency of at least one variable on the remaining covariates, we deduce the covariates to be = 6.  




## Exercise 2

```{r, results=FALSE, include=FALSE}
library(dplyr)
```
```{r}
data %>% count(children)
```

the highest number is 296 (1)

```{r}
max(data$children)
```

However, if the question addresses the highest number of children per single household who is covered by health insurance we find to following answer: the highest n. of children is 5 per 14 household.




## Exercise 3

```{r}
a=levels(data$region)
b=length(a)
s.m=c()
for (i in 1:b){
  c=length(which(data$region==a[i]&data$smoker=="yes"))/length(which(data$region==a[i]))
  
  sss=paste("The share of smokers in", a[i],"is",as.character(c))
  
  print(sss)
  s.m=c(s.m, c)
}
names(s.m)=a
```

Thus, the region with lowest share of smokers is:
```{r}
names(s.m)[s.m==min(s.m)]
```

### Alternative with matrix table:
```{r}
cond_matrix = matrix(NA, nrow=length(levels(data$smoker)), ncol=length(levels(data$region)))
```

This matrix is indeed empty and full of NA.
```{r}
a=levels(data$region)
region_matrix=matrix(a,ncol=length(a))
region_matrix
colnames(cond_matrix)=region_matrix
rownames(cond_matrix)=levels(data$smoker)

for (a in c(1:4)) {
  cond_matrix[1,a] <- sum(data$smoker=="no" & data$region==region_matrix[,a])/sum(data$region==region_matrix[,a])
  cond_matrix[2,a] <- sum(data$smoker=="yes" & data$region==region_matrix[,a])/sum(data$region==region_matrix[,a])
  
}
print(cond_matrix)
```

Proceed with a for loop, print(cond_matrix) is the matrix which has the full answer summed in a table for each region.

```{r}
min(cond_matrix[2,])
```
Northwest is thus the region with the lowest share of smokers; the share is 0.1689655.




## Exercise 4

```{r, results=FALSE}
library(ggplot2)
```
```{r}
ggplot(data, aes(x=age, y=charges, color=smoker)) +
  geom_point(na.rm=TRUE) +
  ggtitle("Scatter Plot of Age vs. Charges Given Smoking Status") + ylab("Charges") + xlab("Age") +
  geom_smooth(method = "lm", se = FALSE)
```

From the scatter plot we find the following patterns: We see that the charges for medical bills for smokers are on average much higher than the charges for people who don't smoke. Having the age variable on the x-axis, we observe that the charges correlate with age. 
As people age, they should expect their medical charges to increase.




## Exercise 5

```{r}
scatter_plot <- function(any_data, x.variable, y.variable, color.variable){
            
      if ( class(any_data)!="data.frame" ) {
    stop("any_data not of data.frame type")
  }    
  
    ggplot(any_data, aes(x=get(x.variable), y=get(y.variable), color=get(color.variable)))+ 
    geom_point(na.rm=TRUE)+ ggtitle("Scatter Plot") +
    labs(x=x.variable,y=y.variable,colour=color.variable)+
    geom_smooth(method = "lm", se = FALSE)

}
scatter_plot(data,"bmi","charges","sex")
```

We observe that as bmi increases, the variance of charges also increases, therefore the data has heteroskedasticity implied in it. Moreover, we observe in general a weakly positive correlation between the bmi and charges, more so for males.A possible explanation could be that males are more likely to afford medical costs. A possible way to improve the analysis, is by also accounting for the wage differences of between individuals.




## Exercise 6

```{r}
bmi_boxplot <- function(data_inp, split_var){
            
     if ( class(data_inp)!="data.frame" ) {
    stop("any_data not of data.frame type")
  }
  
  ggplot(data=data_inp,aes(x =data_inp$bmi, y=get(split_var)))+
    geom_boxplot(outlier.colour="red", outlier.shape=20, outlier.size=2)+
    labs(x="bmi", y=split_var)

}
bmi_boxplot(data,"region")
```

We observe a difference in the distribution of bmi when we account for each region. For instance, the median in southeast is the highest and also has a wider distribution of individuals with ranging bmi. On the contrary, both regions in the north have a smaller data distribution with the least narrow quartiles. For instance, northwest has the most narrow quartiles. 





