---
title: "PS2_Markdown"
author: 
- José Maria de Mattamouros Resende Fonseca de Oliveira (22-602-783)
- Linda Fiorina Odermatt (17-946-310)
- Elena Trevisani (22-620-603) 
- Vasily Zhuravlev (18-502-401) 
date: "2022-12-02"
output: pdf_document
---

## Exercise 1

```{r, results=FALSE}
load("/Users/nenetrevisani/Desktop/student-mat-test.Rdata")
load("/Users/nenetrevisani/Desktop/student-mat-train.Rdata")
```
```{r}
nrow(test)
nrow(train)
```

There are 143 observations in the test set, and 214 observations in the train set.

```{r}
nrow(test)/(nrow(test)+nrow(train)) 
```

The ratio is roughly 40:60% between the test and the train data sets.
The sizes of the two data sets are within the recommended sizes 20-40% for test set and 60-80% for the training set.

## Exercise 2

```{r}
grade_summary=matrix(data=data.frame(col1=mean(train$G3),
                                     col2=min(train$G3),
                                     col3=max(train$G3)))
colnames(grade_summary)=c("G3")
rownames(grade_summary)=c("Mean","Min","Max")

grade_summary
```

Therefore, we find the following in the training data: the average grade is 11.64, the minimum is 4 and the maximum is 19.00.

## Exercise 3
```{r, results=FALSE}
library(ggplot2)
```
```{r}
ggplot(data=train, aes(x=train$G3)) + 
  geom_histogram(binwidth =1 , color = "darkblue", fill = "lightblue") +
  ggtitle("Histogram for Student Grades in Maths, rounded to the closest figure")+
  ylab("Count")+
  xlab("Final Math Grade (0-20)")
```

Notice the data is fairly normally distributed, there are no large outliers for which we need to transform the data.

## Exercise 4

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

Causal and predictive modelling have in common that they both recur to statistical learning
to define a function $\hat f$ so that $Y \approx \hat f(X)$, given any true set $(X,Y)$. The core difference 
lies in their task and their use for the functional form $\hat f(X)$. Causal models aim at finding causal
relations between covariates and dependent variable; in this view it is concerned with the exact form 
of the function considered. It requires an interpretable form of $\hat f(X)$. Predictive modelling is about 
finding a functional form that best predicts/estimates an unobserved dependent variable. The specific form of 
the function therefore plays lesser of a big role. The important is that the value predicted is
reliable/close to the true one.

## Exercise 5

Absences: presence durin classes has been shown to positively impact a student's engagement with the subject studied.
Schoolsup: extra educational school support indeed may affect the knowledge and engagement of a student with his education. 
Studytime: hours spent although may vary depending on background of esch individual may highly impact the final grade.
Higher: a motivated person may very likely achieve higher grades as math is a core subject in many advanced higher education levels. 
Internet: access to it means access to information.


```{r, results=FALSE}
lm.5vars <- lm(data = train, G3 ~ absences + internet +  schoolsup + higher + studytime)
summary(lm.5vars)

lm.fo.5.interact <- lm(data = train, G3 ~ absences + internet +  schoolsup + higher + studytime + (absences + internet +  schoolsup + higher + studytime )^2)
summary(lm.fo.5.interact)
```

```{r}
MSE.5vars.in.sample <- mean((train$G3 - lm.5vars$fitted.values)^2)
print(paste("MSE Model With 5 Variables:", round(MSE.5vars.in.sample, 4))) 

MSE.fo.5.in.sample <- mean((train$G3 - lm.fo.5.interact$fitted.values)^2)
print(paste("MSE Model With Interaction Variables:", round(MSE.fo.5.in.sample, 4))) 
```

Apparently the MSE is lower with more covariates. This is due to the fact that we train our model in-sample and we test it in-sample. The model logically will do a better job at predicting the test score the more covariates we add, as each will add a little more information to what is in the error term.

## Exercise 6

Variables chosen: Fedu, failures, famsup, paid, freetime. 

```{r, results=FALSE}
lm.10vars <- lm(data = train, G3 ~ absences + internet +  schoolsup + 
                  higher + studytime + Fedu + failures + famsup + 
                  paid + freetime)
summary(lm.10vars)

lm.fo.10.interact <- lm(data = train, G3 ~ absences + internet +  schoolsup + 
                          higher + studytime + Fedu + failures + famsup + 
                          paid + freetime + (absences + internet +  schoolsup + 
                                               higher + studytime + Fedu + failures + famsup + 
                                               paid + freetime)^2)
summary(lm.fo.10.interact)
```

```{r}
MSE.10vars.in.sample <- mean((train$G3 - lm.10vars$fitted.values)^2)
print(paste("MSE Model With 10 Variables:", round(MSE.10vars.in.sample, 4))) 

MSE.fo.10.in.sample <- mean((train$G3 - lm.fo.10.interact$fitted.values)^2)
print(paste("MSE Model With 10 and Interaction Variables:", round(MSE.fo.10.in.sample, 4)))

# We specify new data
fitted.1 <- predict(lm.5vars, newdata = test) # We obtain a new object in the environment with exactly 143 entries.

MSE.5vars.out.sample <- mean((test$G3 - fitted.1)^2)
print(paste("MSE Model With 5 Variables:", round(MSE.5vars.out.sample, 4))) 

fitted.2 <- predict(lm.fo.5.interact, newdata = test)

MSE.fo.5.out.sample <- mean((test$G3 - fitted.2)^2)
print(paste("MSE Model 5 Variables With Interaction Variables:", round(MSE.fo.5.out.sample, 4))) 

fitted.3 <- predict(lm.10vars, newdata = test)

MSE.10vars.out.sample <- mean((test$G3 - fitted.3)^2)
print(paste("MSE Model With 10 Variables:", round(MSE.10vars.out.sample, 4))) 

fitted.4 <- predict(lm.fo.10.interact, newdata = test)

MSE.fo.10.out.sample <- mean((test$G3 - fitted.4)^2)
print(paste("MSE Model 10 Variables With Interaction Variables:", round(MSE.fo.10.out.sample, 4))) 
```

Apparently, the functional form with 10 variables fits the best (model 3).

```{r}
MSEout <- data.frame(model = c("Model.5.vars", "Model.5.inter", "Model.10.vars", "Model.10.inter"), MSE.out = c(MSE.5vars.out.sample, MSE.fo.5.out.sample , MSE.10vars.out.sample , MSE.fo.10.out.sample))
MSEin <- data.frame(model = c("Model.5.vars", "Model.5.inter", "Model.10.vars", "Model.10.inter"), MSE.in = c(MSE.5vars.in.sample, MSE.fo.5.in.sample, MSE.10vars.in.sample, MSE.fo.10.in.sample))

combined.data<- cbind(MSEin, MSEout)

ggplot() +
  geom_col(data = MSEout, aes(model, MSE.out), fill="blue", alpha = 0.2) +
  geom_col(data = MSEin, aes(model, MSE.in), fill="orange2", alpha = 0.6) +
  ggtitle("MSE Barplot") + ylab("MSE") + 
  annotate("text", x=3, y=12, label= "In blue: MSEs out-of-sample") +
  annotate("text", x=3, y=11, label= "In orange: MSEs in-sample")
```

### Fit Plot - Train Sample , In-sample fit plot

Specify Outcome Variable
```{r}
G3_train <- as.matrix(train$G3)
G3_test <- as.matrix(test$G3)
```
Create data frame for plotting predictions from models and print a fit plot.

```{r}
model_var <- c(rep("5 variables",length(G3_train)), rep("5 variables and interaction terms",
                                                        length(G3_train)),
               rep("10 variables",length(G3_train)),rep("10 variables and interaction terms",
                                                        length(G3_train)))
observed_G3 <- c(rep(G3_train, 4))
predicted_G3 <- c(fitted(lm.5vars), fitted(lm.fo.5.interact), fitted(lm.10vars),
                  fitted(lm.fo.10.interact))

fit.plot <- data.frame("observed grades" = observed_G3, 
                       "predicted grades" = predicted_G3, "models" = model_var)

ggplot(fit.plot, aes(x=observed_G3, y = predicted_G3, col = model_var)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, color = "black", size = 1) +
  ggtitle("Scatter Plot") + xlab("Predicted Grades") + 
  ylab("Observed Grades") + labs(col="Model") + ylim(0,20) + xlim(0,20)
``` 
Here we observe that model with 10 variables and interaction terms performs best in comparison to other models.

### Fit Plot - Test Sample , Out-of-sample fit plot

Create data frame for plotting predictions from models and plot a fit plot
```{r}
test_model_var <- c(rep("5 variables",length(G3_test)), rep("5 variables and interaction terms",
                                                            length(G3_test)),
               rep("10 variables",length(G3_test)),rep("10 variables and interaction terms",
                                                       length(G3_test)))
test_observed_G3 <- c(rep(G3_test, 4))
test_predicted_G3 <- c(fitted.1, fitted.2, fitted.3, fitted.4)

test_fit.plot <- data.frame("observed grades" = test_observed_G3, 
                            "predicted grades" = test_predicted_G3, "models" = test_model_var)

ggplot(test_fit.plot, aes(x=test_observed_G3, y = test_predicted_G3, col = test_model_var)) +
  geom_point(size = 1.75) +
  geom_abline(intercept = 0, slope = 1, color = "black", size = 1) +
   ggtitle("Scatter Plot") + xlab("Predicted Grades") + 
  ylab("Observed Grades") + labs(col="Model") + ylim(0,20) + xlim(0,20)
``` 
On an out-of-sample fit we can clearly see that model with 10 variables and interaction terms doesn't make the best predictions, quite on the contrary. 
It is rather hard to observe which model performs the best, but from the look of it one could consider either the model with 5 variables or the model with 5 variables and interaction terms. 
However based on out-of-sample MSE we know that model with 10 variables performs the best with the test data.


