---
title: "PS3_Markdown_file"
author: 
- José Maria de Mattamouros Resende Fonseca de Oliveira (22-602-783)
- Linda Fiorina Odermatt (17-946-310)
- Elena Trevisani (22-620-603) 
- Vasily Zhuravlev (18-502-401) 
date: "2022-12-09"
output: pdf_document
---

```{r, results=FALSE}
load("/Users/nenetrevisani/Downloads/browser_2006.RData")
load("/Users/nenetrevisani/Downloads/browser_new.RData")
```

## Exercise 1

```{r}
print(paste("The average online spending of a household in 2006 was", 
            round(mean(browser_2006$spend),digits=2), "USD"))
```            

The average online spending in 2006 is 1959,921 US dollars.

## Exercise 2

### First alternative

```{r}
id_1297=browser_2006[browser_2006$id == 1297,3:ncol(browser_2006)] #extract row
max(id_1297) # calculate the maximum value in % which represents the
#share of the most visited webpage by the household
sum(id_1297) #should add up to 100

t_id_1297=t(id_1297) #transpose the data frame, so that it can be sorted

id_1297_top_website = head(t_id_1297[order(-t_id_1297),],1) 
# the most visited website

print(paste("The most visited webpage by a household with ID 1297 is",
            names(id_1297_top_website)))

#shortcut function which yields the same result in 1 line

apply(browser_2006[browser_2006$id == 1297,-(1:2)], 1, function(x) max(names(which(x == max(id_1297)))))
``` 

### Second alternative

```{r, warning=FALSE,message=FALSE,results=FALSE}
library(dplyr)
```

```{r}
no_id_spend_browser_2006 <- select(browser_2006, -id, -spend)

max <- colnames(no_id_spend_browser_2006)[max.col(no_id_spend_browser_2006,ties.method="first")]

max[1]
```

The household with ID=1297 is most of the time on the "weather.com" webpage.

## Exercise 3

```{r, warning=FALSE,message=FALSE,results=FALSE}
library("glmnet")
```

```{r}
set.seed(09122022) #for future easier comparision 

lasso_model = cv.glmnet(as.matrix(browser_2006[,-c(1,2)]), browser_2006$spend, 
family = "gaussian", nfolds=5, alpha = 1, dfmax=3) #set the model

vector_coef=c(coef(lasso_model, s = lasso_model$lambda[2])[,1]) 
#lambda such that number of non zero coef is 2, wrap into a vector for simplicity

vector_coef[vector_coef!=0] #all non zero coef plus intercept

#print the names of top 2 websites that are best linear predictors for online spending

top_2=t(vector_coef[vector_coef!=0]) 
names(top_2[,-1])
```

The two webpages that are together the best linear predictors for online
spending in 2006 are "officedepot.com" and "staples.com".

## Exercise 4

```{r, warning=FALSE,message=FALSE,results=FALSE}
library(ggplot2)
```

```{r, warning=FALSE}
#Post-lasso model estimated with OLS method
post.lasso_model = lm(spend ~ (officedepot.com + staples.com), data = browser_2006)

#OLS model with all variables
all.variables_ols = lm(spend ~ . , data = browser_2006[,-1]) 

#Predicted Values for new data
prediction.post.lasso_model = predict(post.lasso_model, newdata = browser_new)
prediction.all.varibales_OLS = predict(all.variables_ols, newdata = browser_new)

#Correlation between OLS and Post-Lasso predictions
cor(prediction.all.varibales_OLS, prediction.post.lasso_model) 

#Plot of fitted values 
predicted.values =as.data.frame(cbind(prediction.all.varibales_OLS,prediction.post.lasso_model),
colnames=c("all.variables_ols", "post.lasso_model")) 

ggplot(data = predicted.values, 
aes(x = prediction.all.varibales_OLS, y= prediction.post.lasso_model)) +
geom_point() + 
geom_abline(intercept = 0, slope = 1) +
ggtitle("Scatter Plot with line of best fit with predicted values for test data")+
ylab("Post-Lasso predicted spending") + 
xlab("OLS predicted spending")
      
#Zoomed in plot without some outliers for better visuals
ggplot(data = predicted.values, 
aes(x = prediction.all.varibales_OLS, y= prediction.post.lasso_model)) +
geom_point() + 
geom_abline(intercept = 0, slope = 1) +
xlim(-10000,15000) + 
ylim(1500,10000) +
ggtitle("Zoomed Scatter Plot with line of best fit")+
ylab("Post-Lasso predicted spending") + 
xlab("OLS predicted spending")
```

conclusion missing - observing the graph it seems lasso provides more consistent
and lower variance predictions - need to justify with/ wether lasso is better

From the surface, it seems that a post-Lasso model should create more reasonable
predictions based on the fact that it generates only positive values. On the other
hand if OLS based prediction is used then we can see that it can become problematic
as it would be possible to generate negative predicted values. In essence, it
doesn't make sense since spending extra time on a website won't result in customers
receiving the money for online spending.

Moreover, judging from the in-sample R-squared values, post-Lasso Model doesn't
do a great job in explaining the variability in data and therefore it impacts
it may impact its ability to predict on out-of-sample data. In our view,
it makes sense to include more control variables in order to improve the
predictability of post-Lasso model.



